Methods for Large Scale Learning to Rank

### 2014-09-17 report/Chapters/Implementation.tex
wat refactoring + werkverdeling in thesis

Microsoft offers a scalable and on-demand Hadoop service with HDInsight, which enables Hadoop services for those not able to make the required investments for their own cluster. The latest HDInsight version available at the time of writing, HDInsight 3.1, runs the Hortonworks distribution of Hadoop version 2.4. While early versions of Hadoop were merely an open source implementation of MapReduce, newer versions since Hadoop 2.0 offer support for variety of programming models with the introduction of Hadoop YARN \cite{Vavilapalli2013}. Newly supported programming models since Hadoop 2.0 include Dryad \cite{Isard2007}, Giraph \cite{Avery2011}, MPI, REEF \cite{Chun2013}, Spark \cite{Zaharia2010}, and Storm \cite{Aniello2013}. Even though these programming models are now supported by Hadoop and even though some of these programming models have recently increased in popularity, they still lack the critical mass as the data processing framework of choice that MapReduce has as Lin argued back in 2012 \cite{Lin2013}(see section \ref{sec:motivation}). Therefore, even though some of these programming models might be a better fit for iterative algorithms, we use Hadoop MapReduce of the programming model to implement the Learning to Rank algorithms.\\

@@

>>>>

Microsoft offers a scalable and on-demand Hadoop service with HDInsight, which enables Hadoop services for those not able to make the required investments for their own cluster. The latest HDInsight version available at the time of writing, HDInsight 3.1, runs the Hortonworks distribution of Hadoop version 2.4. While early versions of Hadoop were merely an open source implementation of MapReduce, newer versions since Hadoop 2.0 offer support for variety of programming models with the introduction of Hadoop YARN \cite{Vavilapalli2013}. HDInsight Hadoop data nodes are regular Large A3 Microsoft Azure virtual machines with four core processors and 7GB RAM. Newly supported programming models since Hadoop 2.0 include Dryad \cite{Isard2007}, Giraph \cite{Avery2011}, MPI, REEF \cite{Chun2013}, Spark \cite{Zaharia2010}, and Storm \cite{Aniello2013}. Even though these programming models are now supported by Hadoop and even though some of these programming models have recently increased in popularity, they still lack the critical mass as the data processing framework of choice that MapReduce has as Lin argued back in 2012 \cite{Lin2013}(see section \ref{sec:motivation}). Therefore, even though some of these programming models might be a better fit for iterative algorithms, we use Hadoop MapReduce of the programming model to implement the Learning to Rank algorithms.\\

@@

We create a utility class HDInsightWrapper, which abstracts away connection setup, authorisation and response parsing from the tasks of submitting jobs to the cluster and retrieving data from \ac{WASB} storage. The HDInsightWrapper class is parametrised with connection details like the cluster name, cluster user, cluster password, storage account, storage container and storage key, and has two methods:

\begin{description}
\item[void runPig(String pigCode)]{Receives a Pig script in String format as input and handles the submission of this script as MapReduce job on the cluster. Polls the Templeton REST API with progress requests every 50 milliseconds until the MapReduce job has completed. Blocks until the MapReduce job has completed.}
\item[String runPig(String pigCode, String outputDir)]{Identical to runPig(String pigCode), but reads the result of the computation from location 'outputDir' after completion of the MapReduce job and outputs the result as String value.}
\end{description}

Setting the degree of parallelisation is crucial to obtain optimal cluster utilisation. HDInsight cluster machines offer four mapper slots and two reducer slots per data node. Cluster utilisation is optimal when the computational work is evenly distributed over exactly the number of available mappers in the map phase, and evenly distributed over exactly the number of reducers in the reduce phase. The number of mappers used can be controlled by setting MapReduce configuration parameters mapreduce.input.fileinputformat.split.maxsize, mapreduce.input.fileinputformat.split.minsize, and the Pig configuration parameter pig.maxCombinedSplitSize to the data size in bytes of the input training data divided by the number of available mappers. To control the number of used reducers, Pig offers two options: 1) the default\_parallel parameters, which can set a default number of reducers used throughout all MapReduce jobs that the Pig script consists of, and 2) the "PARALLEL" clause, which sets the number of reducers at operator level (overrides default\_parallel if set). The framework uses listed configuratoin parameters to control the numbers of mappers used and uses the "PARALLEL" clause to set the number of used reducers.
